Generality in food webs scales with species richness, not latitude

### 2014-07-11 manuscript/latitudepaper.tex
Forget what I changed.

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs (networks of feeding links between species) have been used for decades to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, temperature?, latitude?. 
Latitude is of particular interest as an environmental driver of food-web structure because it entails gradients in 
temperature \citep{}, primary productivity \citep{}, and seasonality \citep{}. Importantly, there is also a strong
latitudinal gradient in species richness for most taxa \citep{Schemske2009a}.

This effect of latitude on species richness, however, presents a problem for the analysis of food web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that relationships
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties that both scale with species richness and are related to candidate
explanations for the latitudinal diversity gradient, such as measures of specialization.

@@

Importantly, the expectation of the latitude niche-breadth hypothesis that higher species richness is associated with 
reduced competition is contary to the expected increase in the mean number of interactions per species (one potential 
measure of specialization) as species richness increases \citep{Dunne2006,Riede2010}. If, therefore, there is a
distinct effect of latitude on specialization then this effect should be visible in changing scaling relationships.
Here, we tested the effect of latitude on scaling relationships between species richness and mean links per species, 
mean generality (number of prey per predator), and mean vulnerability (number of predators per prey). Together, these
properties provide a measure a species' biotic specialization - that is, its specialization in terms of interactions 
with other species.

@@

\subsection*{Data Set}
We compiled a list of XX empirical food webs from an online database (GlobalWeb webs \citep{GlobalWeb}), supplemented 
with XX food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
Of the GlobalWeb webs, XX were rejected because their original source could not be found (XX), they were source or sink
webs rather than descriptions of an entire community (XX), were focused on plant-pollinator, host-parasitoid, or 
competitive interactions (XX), described inferred interactions in an extinct community (2), were ``generalized 
schemes'' rather than being based on empirical observation (5), or because it was not clear which of a variety of 
described sites the published food web represented (1). (See Table S1 for a list of rejected webs and reasons for their
exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
interactions between them \citep{Dunne2013}. For each of the XXX webs included in the final database, we recorded the
study site latitude if possible or, where no latitude was given, obtained estimated coordinates of the described
study site using Google Earth \citep{GoogleEarth}. Webs were also classified according to ecosystem type (terrestrial,
lake, stream, estuary, or marine), year of publication (to control for changing standards of data collection over time \citep{Dunne2006}), and whether or not humans were included (11 webs included humans).

@@

\begin{equation}
\label{PowerLat}

Z_{i}=\alpha S^{\beta_{0}+\beta_{1}\Delta_{i}+\beta_{3}\Sigma_{i} + \beta_{4}\Lambda_{i} + \beta_{5}T_{i} + \beta_{6}M_{i} 
+ \beta_{7}\Delta_{i}\Sigma_{i} + \beta_{8}\Delta_{i}\Lambda_{i} + \beta_{9}\Delta_{i}T_{i} + \beta_{10}\Delta_{i}M_{i}}
+ \rho{1}\Delta_{i} + \rho_{2}\Upsilon_{i} + \rho_{3}H_{i} + P_{i} + \epsilon_{i} 
\end{equation}

where $\delta$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$\Sigma_{i}$, $\Lambda_{i}$, $T_{i}$, and $M_{i}$ are 
factors indicating whether web $i$ is a stream, lake, 
terrestrial, or marine food web (e.g., $\Sigma_{i}$=1 for 
a stream food web, 0 otherwise). If $\Sigma_{i}$, $\Lambda_{i}$, $T_{i}$, and $M_{i}$ are all 0, web $i$ 
describes an estuary food web. $\Upsilon_{i}$ is the year web $i$ was published, $H_{i}=1$ if web $i$ contains humans, 0 otherwise, and $P_{i}$ is a random effect of 
site ID. Note that all fixed effects $\beta$ refer to factors acting on the scaling exponent between $Z$ and $S$
while fixed effects $\rho$ refer to factors acting on the residual error.

@@

\begin{equation}
\label{LogLat}
Z_{i}= log(\alpha)+\beta_{0}log(S) + \beta_{1}\Delta_{i}log(S) +\beta_{3}\Sigma_{i}log(S) + \beta_{4}\Lambda_{i}log(S)  + \beta_{5}T_{i}log(S)  + \beta_{6}M_{i}log(S)  
+ \beta_{7}\Delta_{i}\Sigma_{i}log(S)  + \beta_{8}\Delta_{i}\Lambda_{i}log(S)  + \beta_{9}\Delta_{i}T_{i}log(S)  + \beta_{10}\Delta_{i}M_{i}log(S) 
+ \rho_{1}\Delta_{i} + \rho_{2}\Upsilon_{i}) + \rho_{3}H_{i} +\epsilon_{i}

\end{equation}

@@

\begin{tabular}

\label{Latlms}
\caption{Intercepts and slopes for simple linear regressions of food web properties against absolute latitude.}
\begin{table}{l | l l l l}
\multirow{2}Property & \multicolumn{2}{c}{Intercept} & \multicolumn{2}{c}{Slope} \\
 & $\alpha$ & $p$-value & $\alpha$ & $p$-value \\
\hline
$S$ & 23.10 & \textless0.001 & 0.284 & 0.040 \\
$Z$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
$G$ & 3.269 & \textless0.001 & 0.049 & 0.014 \\
$V$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
\end{table}
\end{tabular}

@@

\begin{tabular}

\label{Errorfits}
\caption{AIC's for model \ref{Power}, assuming normal error, and model \ref{Loglog}, assuming lognormal error.}
\begin{table}{l | l l | c}
Property & Normal error & Lognormal error & Model selected \\
\hline
$Z$ & 1151 &  767 & \ref{Loglog} \\
$G$ & 1266 & 1037 & \ref{Loglog} \\
$V$ & 1151 & 767 & \ref{Loglog} \\
\end{table}
\end{tabular}

In this data set, none of the best fit models for species richness, connectance, or links per species included an 
effect of latitude on either the scaling exponent or residual (Table \ref{bestfits}). 
$G$ increased slightly in more recent webs ($\rho{2}$=0.005, $p$<0.001), but the best-fit models for $V$ and $Z$ did not include a term for year of publication. 
In contrast, $Z$ and $V$ were higher in lake food webs ($\beta_{4}$=0.10, $p$<0.001 and $\beta_{4}$=0.10, $p$<0.001, respectively) while the best-fit model for $G$ did not include an effect for any ecosystem type.

>>>>

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs (networks of feeding links between species) have been used for decades to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, temperature?, latitude?. 
Latitude is of particular interest as an environmental driver of food-web structure because it entails gradients in 
temperature \citep{}, primary productivity \citep{}, and seasonality \citep{}. Indeed, each of these latitudinal 
environmental gradients has been put forward as a driver for the latitudinal gradient in species richness, one of
the most general and robust patterns in ecology \citep{Schemske2009a,,}.

This effect of latitude on species richness, however, presents a problem for the analysis of other aspects of food-web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that relationships
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties that both scale with species richness and are related to candidate
explanations for the latitudinal diversity gradient, such as measures of specialization.

@@

Importantly, the expectation of the latitude-niche breadth hypothesis that higher species richness is associated with 
reduced competition is contary to the expected increase in the mean number of interactions per species (one potential 
measure of specialization) as species richness increases \citep{Dunne2006,Riede2010}. If, therefore, there is a
distinct effect of latitude on specialization then this effect should be visible in changing scaling relationships.
Here, we tested the effect of latitude on scaling relationships between species richness and mean links per species, 
mean generality (number of prey per predator), and mean vulnerability (number of predators per prey). Together, these
properties provide a measure a species' biotic specialization - that is, its specialization in terms of interactions 
with other species.

@@

\subsection*{Data Set}
We compiled a list of 347 empirical food webs from an online database (GlobalWeb webs \citep{GlobalWeb}), supplemented 
with 31 food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
Of the GlobalWeb webs, 122 were rejected because their original source could not be found or was unpublished (53), they were source or sink
webs rather than descriptions of an entire community (13), were focused on plant-pollinator, host-parasitoid, or 
competitive interactions (8+34+2=44), described inferred interactions in an extinct community (2), were ``generalized 
schemes'' rather than being based on empirical observation (8), or because it was not clear which of a variety of 
described sites the published food web represented (2). (See Table S1 for a list of rejected webs and reasons for their
exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
interactions between them \citep{Dunne2013}. For each of the 263 webs included in the final database, we recorded the
study site latitude if possible or, where no latitude was given, obtained estimated coordinates of the described
study site using Google Earth \citep{GoogleEarth}. Webs were also classified according to ecosystem type (terrestrial,
lake, stream, estuary, or marine), year of publication (to control for changing standards of data collection over time \citep{Dunne2006}), and whether or not humans were included (11 webs included humans).

@@

\begin{equation}
\label{PowerLat}
Z_{i}=\alpha S^{\beta_{0}+\beta_{1}\Delta_{i}+\beta_{3}\Sigma_{i} + \beta_{4}\Lambda_{i} + \beta_{5}T_{i} + \beta_{6}M_{i} 
+ \beta_{7}\Delta_{i}\Sigma_{i} + \beta_{8}\Delta_{i}\Lambda_{i} + \beta_{9}\Delta_{i}T_{i} + \beta_{10}\Delta_{i}M_{i}}
+ \rho{1}\Delta_{i} + \rho_{2}\Upsilon_{i} + \rho_{3}H_{i} + P_{i} + \epsilon_{i} 
\end{equation}

where $\delta$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$\Sigma_{i}$, $\Lambda_{i}$, $T_{i}$, and $M_{i}$ are 
factors indicating whether web $i$ is a stream, lake, 
terrestrial, or marine food web respectively (e.g., $\Sigma_{i}$=1 for 
a stream food web, 0 otherwise). If $\Sigma_{i}$, $\Lambda_{i}$, $T_{i}$, and $M_{i}$ are all 0, web $i$ 
describes an estuary food web. $\Upsilon_{i}$ is the year web $i$ was published, $H_{i}=1$ if web $i$ contains humans, 0 otherwise, and $P_{i}$ is a random effect of 
site ID. Note that all fixed effects $\beta$ refer to factors acting on the scaling exponent between $Z$ and $S$
while fixed effects $\rho$ refer to factors acting on the residual error.

@@

\begin{equation}
\label{LogLat}
Z_{i}= log(\alpha)+\beta_{0}log(S) + \beta_{1}\Delta_{i}log(S) +\beta_{3}\Sigma_{i}log(S) + \beta_{4}\Lambda_{i}log(S)  + \beta_{5}T_{i}log(S)  + \beta_{6}M_{i}log(S)  
+ \beta_{7}\Delta_{i}\Sigma_{i}log(S)  + \beta_{8}\Delta_{i}\Lambda_{i}log(S)  + \beta_{9}\Delta_{i}T_{i}log(S)  + \beta_{10}\Delta_{i}M_{i}log(S) 
+ \rho_{1}\Delta_{i} + \rho_{2}\Upsilon_{i}) + \rho_{3}H_{i} +\epsilon_{i}
\end{equation}

@@

\begin{table}[!h]
\caption{Intercepts and slopes for simple linear regressions of food web properties against absolute latitude.}
\label{Latlms}
\begin{tabular}{l | l l | l l}

\multirow{2}{*}Property & \multicolumn{2}{|c}{Intercept} & \multicolumn{2}{|c}{Slope} \\
                      & $\alpha$ & $p$-value & $\alpha$ & $p$-value \\
\hline
$S$ & 23.10 & \textless0.001 & 0.284 & 0.040 \\
$Z$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
$G$ & 3.269 & \textless0.001 & 0.049 & 0.014 \\
$V$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
\end{tabular}
\end{table}

@@

\begin{table}[!t]

\caption{AIC's for model \ref{Power}, assuming normal error, and model \ref{Loglog}, assuming lognormal error.}
\label{Errorfits}
\begin{tabular}{l | l l | c}
Property & Normal error & Lognormal error & Model selected \\
\hline
$Z$ & 1151 &  767 & \ref{Loglog} \\
$G$ & 1266 & 1037 & \ref{Loglog} \\
$V$ & 1151 & 767 & \ref{Loglog} \\
\end{tabular}
\end{table}

In this data set, none of the best fit models for species richness, connectance, or links per species included an 
effect of latitude on either the scaling exponent or residual (Table \ref{Bestfits}). 
$G$ increased slightly in more recent webs ($\rho_{2}$=0.005, $p$\textless0.001), but the best-fit models for $V$ and $Z$ did not include a term for year of publication. 
In contrast, $Z$ and $V$ were higher in lake food webs ($\beta_{4}$=0.10, $p$\textless0.001 and $\beta_{4}$=0.10, $p$\textless0.001, respectively) while the best-fit model for $G$ did not include an effect for any ecosystem type.

\begin{table}[!h]

\caption{Standard effects and their $p$-values for the best-fitting models of $Z$, $G$, and $V$.}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l | l l }
\multirow{2}{*}{Property} & \multicolumn{2}{c}{Intercept} & \multicolumn{2}{c}{log(Species)} & \multicolumn{2}{c}{log(Species)$\Lambda$} & \multicolumn{2}{c}{$\Upsilon$} \\
& $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value  \\
\hline
$Z$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
$G$ & -9.999 & \textless0.001 & 0.547 & \textless0.001 & \multicolumn{2}{c}{NA} & 0.005 & \textless0.001 \\
$V$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
\end{tabular}
\end{table}
